---
title: "Dana"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load tidyverse
```{r}
library(tidyverse)
```
load data

```{r}
race_results <- read_csv("race_results_statsf1.csv")
```

decide on questions:

1. what teams have seen the most successful drivers?
1. which team have raced the most F1 races?

the story will possibly be about if the best/most successful F1 teams are the legacy teams or have newer competitors managed to gain any advantage.

```{r}
summary(race_results)
```
## Question 2

which team have raced the most F1 races?

```{r}
competed<-filter(race_results, position != "f")
just_teams <- competed %>% select(race_id, team) %>% distinct(race_id,team)
```

We can count this way:
```{r}
competed_teams <- just_teams %>%
                  group_by(team) %>%
                  summarize(number_of_races = n())
```

These teams competed. We want to count how many races each team has. Let's group by teams
```{r}
competed_teams <- group_by(competed,team) 
```
Now that we grouped them. Let's count how many races they each had.

```{r}
summarize(competed_teams, number_of_races=n_distinct())
```

